---
title: "Handling Imaging-MS data with the MsBackendImzML"
output:
    BiocStyle::html_document:
        toc_float: true   
vignette: >
  %\VignetteIndexEntry{MsBackendImzML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

**Package**: `r Biocpkg("MsBackendImzML")`<br />
**Authors**: `r packageDescription("MsBackendImzML")[["Author"]] `<br />
**Last modified:** `r file.info("MsBackendImzML.Rmd")$mtime`<br />
**Compiled**: `r date()`

```{r, echo = FALSE, message = FALSE}
library(MsBackendImzML)
library(BiocStyle)
```


# Reading the imzML files

When loading the imzML files, please make sure that the corresponding ibd file
is also present on the same folder as the imzML.

```{r}
file_path <- system.file("/extdata/test_image.imzML", package = "MsBackendImzML")
be <- backendInitialize(MsBackendImzML(), files = file_path)
sp <- Spectra(be)
```

As you can see, this backend has some extra `spectraVariables` in order to
handle the pixel positions and the byte offsets corresponding to the ibd
(imaging binary data) file:

```{r}
spectraVariables(sp)
```
# Exploring the data

As with all `Spectra` objects, we can extract the `peaksData()` (mz, intensity and, if
available, inv_ion_mobility). The data will then be read on-the-fly from the ibd
file/s and loaded into R.

```{r}
pd <- peaksData(sp)
print(head(pd[[1]], n=21))
```


```{r}
plot(head(pd[[1]], n=7), type = "b", col=2)
```

As you can probably infer from the plot, this data is in profile mode: the
MsBackendImzML does **not** make any assumptions (nor checks) about the data
origin or whether any processing steps (smoothing, alignment,
centroiding/peak-picking, ...) have been applied to the data. Therefore, it is
tolerant to continuous/processed mode data as well as raw/profile/centroid data.

# Session Info

```{r}
sessionInfo()
```


